---
title: Introducing R (to someone)
date: '2018-03-07'
slug: intro-to-r
categories:
  - rbloggers
tags: ["R","tidyverse", "rstats101"]
output:
  blogdown::html_page:
    toc: true
---

```{r include=FALSE}
knitr::opts_chunk$set(comment = "#>", collapse = TRUE)
```

# May I introduce you...
*...to my favourite programming language?*

every once in a while you want to introduce someone new to the tidyverse. But how do you do it and where do you start?

(you have the chance to make the first impression.)

- setting: at work
- show that R can produce awesome results easily (get to the quirks later...)
- show where to learn more (cheatsheets)


- look for some of the cool resources to learn R

## some helpful resources
Here are some resources I've found:

* http://paldhous.github.io/NICAR/2018/r-analysis.html (recommended by [\@PatrickStotz](https://twitter.com/PatrickStotz))
* https://github.com/hadley/building-permits (from -link to video)
* http://stat545.com/topics.html (by JennyBC)
* http://r4ds.had.co.nz (classic!)


## PS
Thank you, R.️

```{r rstats love, echo=FALSE}
blogdown::shortcode('tweet', '970314363696570368')
```